The impacts of park access on health outcomes: a spatial comprehensive approach

### 2019-01-05 manuscript.Rnw
Add literature review from Nico, revised and expand

The United States is currently facing an epidemic of obesity and chronic
diseases, which are non-communicable diseases of long duration and typically
slow development, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer. The \citet{WHO2014} estimates approximately 300,000
premature deaths in the United States each year can be attributed to chronic diseases.
\citet{Finkelstein2009} find an
obese American spends nearly \$1,500 more per year in health care costs than an
American of normal weight. As obesity and chronic disease
have become rampant, it is no surprise that healthcare costs have risen to
nearly one-fifth of the United States’ gross domestic product \citep{Harnik2011}.

While a moderate amount of regular physical activity has been established as an
effective strategy for reducing and managing obesity and many of the
aforementioned chronic diseases \citep{Gordon2013}.
\citeauthor{Wolf2008} notes that “more than 50 percent of U.S. adults do not get
enough physical activity to provide health benefits; 24 percent are not active
at all in their leisure time. Activity decreases with age and sufficient
activity is less common among women than men, and among those with lower incomes
and less education” \cite[22]{Wolf2008}. Similarly, the United States Department of
Health and Human Services notes that less than fifty-percent of Americans meet
established recommendations for moderate to vigorous physical activity, or MVPA
(U.S. DHHS 2010). As the epidemiological transition from infectious to chronic
diseases is now complete in the developed world, increasing physical activity
has become a vital public health task in the 21st century.

Research has shown that the design of the built environment influences a range
of behaviors, including those related to physical activity. The presence of
trees and other vegetation in outdoor environments have been shown to be
positively associated with physical activity \citep{Pretty2005}; one finding
suggests that after sidewalks and trails have been constructed, the introduction
of natural elements positively impacts motivation to engage in physical activity
\citep{Suminski2005}. Additional evidence indicates that commonly vegetated
areas, such as parks and open space, support outdoor physical activity
\cite{Giles-Corti2005, Wells2007}. Perhaps most telling is the
finding that “people in large cities perceive themselves to be generally
healthier if a greater percentage of the living environment is greenspace, are
inclined to be more active, and claim the ability to relax faster”
\cite[24]{Wolf2008}. Thus, by providing space for active recreation, public
parks and greenspaces may increase the likelihood of engaging in more physical
activity. As such, public investment in parks can be thought of as a public
health intervention for chronic diseases and conditions, as well as general
population health.

To date, the literature exploring the relationship between parks and health
outcomes, specifically those related to physical activity and obesity, has
yielded mixed results. Importantly, there is considerable variation in the
design of past studies, including the spatial scale of analysis, the population
of interest, and the measure of proximity and/or accessibility. While studies of
neighborhood level health impacts do exist, they typically focus on discrete
populations of interest within individual parks instead of examining the impacts
of larger park networks \citep{Bancroft2015, Coutts2008}.
Other studies have used the city and metropolitan statistical area as the
spatial scale of analysis, which
in some cases has resulted in the discovery of positive associations between
increased park space and positive health impacts but may conceal more nuanced
relationships that exist at smaller spatial scales \citep{Larson2016}.
Evaluating the health impacts of parks and greenspaces using a smaller
unit of analysis is an important research gap to fill, especially considering
that park use and physical activity within parks varies considerably according
to residential proximity to parks and park facilities as well as a number of
sociodemographic factors \citep{Kaczynski2014}. There are important urban
design considerations here; is it better to build a single large park with many
different amenities, or to build a series of smaller parks nearer people’s
homes. Similarly, do parks placed near low-income households affect the behavior
of those households? This study attempts to fill that gap using New York City as
a case study example. Using a new measure of park access called ‘Park Choice
Accessibility,’ we find

\section{Existing Literature}

\subsection{Urbanization and Public Health Outcomes}
Across the world, the population of cities is growing at an unprecedented rate.
With a current estimated global urban population of over 3.5 billion people, the
United Nations Population Fund estimates that by 2030, 5 billion people will
inhabit cities worldwide.  By 2050, an additional 3 billion people will live in
cities – a sixty percent increase over twenty years – with much of this growth
expected to take place in the developing world (United Nations Population Fund,
2016). Despite the greater expected concentration of urban growth in other parts
of the globe, the trend of urbanization is highly visible in the United States
as well.  According to the U.S. Census Bureau, ''The nation's urban population
increased by 12.1 percent from 2000 to 2010, outpacing the nation's overall
growth rate of 9.7 percent for the same period.'' \citet{Nowak2005}
 estimate that urban land area as a percentage of total land area in the
U.S. will increase from 3.1\% in 2000 to approximately 8.1\% by the year 2050,
with urbanized places collectively comprising a land area larger than the state
of Montana. More than eighty percent of the U.S. population now lives in
urban areas, compared to sixty-four percent in 1950 (United States Census Bureau
2007).

The global trend of urbanization has profound implications for population
health, both positive and negative. Because urbanization corresponds with an
increase in population density compared to rural and suburban settlement
patterns, residents of dense urban areas throughout the world have better access
on average to many health services and health-promoting amenities simply because
of their closer proximity to such resources \citep{Larson2016}.
However, despite this benefit of increased density and the ability of cities to
provide many opportunities for innovation, economic growth and social progress,
historical and current evidence suggests that rapid urban growth often leads to
congestion and numerous negative environmental and human health outcomes.
Interactions between growing urban populations and their environment, marked by
intensive resource consumption and ecologically harmful patterns of development,
lead to numerous undesired consequences including pollution and sanitation
issues as well as racial and socioeconomic disparities \citep{Larson2016}.

Because of the considerable growth of urbanization throughout the world, many
city planning and public health professionals have begun to pay more attention
to the role of the built environment in promoting or discouraging health
behaviors. Until recently, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about health
lifestyles and behaviors, touching on topics including diet and exercise.
According to \citet{Coutts2008} however, this trend is shifting as professionals
''have begun adopting an ecological paradigm, accepting that both individual and
environmental determinants play a role in health behavior. This new, arguably
revisited, public-health paradigm accounts not only for the compositional (who
you are) but also for the contextual (where you are) influences on physical
activity.'' As professionals begin to operate from the assumption that the design
and configuration of the built environment can facilitate or inhibit physical
activity, they are increasingly looking to public spaces like parks and
greenways as key elements of the built environment that can support exercise
\citet{Coutts2008, Bedimo-Rung2005}.

\subsection{Review of the Health Benefits of Parks}
A large body of literature exists documenting the many benefits of urban parks
and greenspaces to human health and wellbeing. Importantly, these benefits stem
from the provision of ecosystem services, which occur when the natural
environment supplies something that people demand, improving quality of life and
well-being \citep{Larson2016}. These services can include the
provisioning of goods such as fresh water and agricultural products; regulatory
functions including protection of drinking water quality, heat mitigation, air
purification, and stormwater management; and cultural functions, such as
improving aesthetics, providing opportunities for recreation, tourism, and
physical and mental health, and promoting biodiversity \citep{McDonald2015}.
According to \citet{Wolch2014}, parks and greenspaces provide ''\ldots a wide
range of ecosystem services that could help combat many urban ills and improve
life for city dwellers—especially their health…Ecosystem services provided by
urban greenspace not only support the ecological integrity of cities, but can
also protect the public health of urban populations.'' Table
\ref{tab:parkbenefits} summarizes commonly cited benefits of parks and
greenspaces.

\begin{table*}[ht]
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{11cm}}
\toprule
Category & Summary of Benefits\\
\midrule
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Foster increased physical activity\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce of heat-related mortality \\
 & Reduce incidence of cardiovascular-related mortality \\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
 & Reduce incidence of childhood asthma \citep{Wolf2008}\\
 & Improve birth outcomes\\
\midrule
Mental Health & Reduce stress and mental fatigue\\
  & Reduce aggression\\
  & Enhance emotional and cognitive development\\
  & Improve behavioral outcomes in youth \\
\midrule
Social Health & Enhanced community aesthetics\\
 & Reduce crime\\
 & Increase social interaction \\
\midrule
Economic Health & Provide of ecosystem services\\
 & Increase residential property values and municipal property tax revenues\\
 & Attract more shoppers and increase economic activity to commercial districts \\
\bottomrule
\end{tabular}
\end{table*}

Given the general benefits of parks and greenspaces in Table \ref{tab:parkbenefits}, several studies
have attempted to quantify the impacts of these spaces on different facets of
health and wellbeing at the city level, yielding mixed results. \citet{Larson2016}
used self-reported scores on the Gallup-Healthways Wellbeing
Index to evaluate the relationship between different areas of wellbeing,
including physical health, and park quantity, quality, and accessibility in 44
U.S. cities. The authors found that “Park quantity (measured as the percentage
of city area covered by public parks) was among the strongest predictors of
overall wellbeing, and the strength of this relationship appeared to be driven
by parks’ contributions to physical and community wellbeing” (Larson, Jennings \&
Cloutier 2016). While the authors found positive associations between wellbeing
and park quality and accessibility, these relationships were not statistically
significant. A study by \citet{West2012} used park data from the
Trust for Public Land’s 2010 City Park Facts and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
between the density of parkland, parkland per capita, and levels of physical
activity and obesity for 67 metropolitan statistical areas in the U.S. The study
found a significant, positive association between park density and physical
activity and a significant, negative association between park density and
obesity. In a study of New York City, \citet{Stark2014} found that the
“proportion of neighborhoods that was large or small park space and park
cleanliness were associated with lower BMI among NYC adults after adjusting for
other neighborhood features such as homicides and walkability, characteristics
that could influence park usage.”

Interestingly, in a study by \citet{Richardson2012} that examined the
relationship between urban greenspace and selected mortality rates, the authors
did not find a statistically significant relationship between the quantity of
urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
car accidents, but found that all-cause mortality was significantly higher in
greener cities. In a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured physical
activity, \citet{Bancroft2015} found that five studies reported a significant
positive association between the two, six studies produced mixed results, and
nine studies found no association at all.

While several of the aforementioned studies yielded results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, all of the
studies  -- except some that were included in the meta-analysis by
\citet{Bancroft2015} -– were conducted at the level of the city or metropolitan statistical
area, potentially concealing variation in health outcomes at more fine-grained
levels of analysis \citep{Larson2016}.

\subsection{Accessibility to Parks}
To evaluate the impacts of parks and greenspaces on health outcomes at any
scale, one of the fundamental tasks involved is the selection of a measure of
accessibility. In a comprehensive meta-analysis of published studies that
measure active accessibility (accessibility using non-motorized travel modes
including walking and cycling), \citet{Vale2016} identified four
broad categories of studies based on the active accessibility metric employed:
“distance-based, gravity-based or potential, topological or
infrastructure-based, and walkability and walk score-type measures.”
Distance-based measures account only for the Euclidean distance between origins
and destinations, while infrastructure-based measures explicitly incorporate
relevant transportation networks like roads and sidewalks to more accurately
measure travel time and distance. Gravity-based measures incorporate cost
measures to model accessibility as a function of a destination’s attractiveness
(i.e., size, commercial activity, etc.) and the cost of traveling to that
destination from a given origin (i.e., travel time or distance).

Importantly, \citeauthor{Vale2016} acknowledge that there is not yet a
consensus on the most appropriate accessibility metric to use in a given
setting, noting that ''ways to measure active accessibility are as varied as the
number of scholars the measure them.'' However, their meta-analysis
highlights some of the limitations associated with each of the types of metrics
described. For example, in describing distance-based accessibility measures, the
authors note that such measures are ''extremely sensitive to the way in which
travel impedance (i.e., distance) is measured. Accordingly, four types of
distance can be identified: Euclidean distance, Manhattan distance, shortest
network distance, and shortest network time''. The
appropriateness of one distance-based measure over another can vary
significantly depending on the topography of the environment and the travel mode
that is being employed. In describing gravity-based accessibility measures,
which ''assume that travel is a derived demand and there is a tradeoff between
the benefit of the opportunity and the cost to reach it from a given origin,''
the authors note that such measures do not always explicitly account for land
use characteristics near origins and destinations, which may impact that true
accessibility of those places.

In evaluating methodologies of measuring access to urban services, including
parks and greenspaces, \citet{Logan2017} note that existing and commonly used
approaches “often simplify their measure of proximity by using large areal
units and by imposing arbitrary distance thresholds,” which often results in
access-poor populations being overlooked. The authors concede that many
existing approaches have long been necessary because the computational power
required to use higher-resolution analytical techniques was unavailable.
However, due to recent advances in computation power and the advent of
municipal open-data policies, \citeauthor{Logan2017} recommend that future
analyses of accessibility disaggregate population data to the building or
parcel level and use network distance instead of Euclidean distance to measure
proximity.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus, or the total value of all alternatives of the choice set, can 
be obtained as the log-sum of the denominator of the choice probability equation:
\begin{equation}\label{eq:cs}
CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher. This allows for the fact that some 
people are more or less sensitive to distances, and that distance is a 
continuous, and not a binary, phenomenon. Second, the random utility 
formulation allows the researcher to include any attribute of the park; 
in this case, we consider the
size of the park as an element of accessibility. This suggests that not all
parks are equal, and that a large park such as Central Park on Manhattan may
provide health and activity benefits over a much larger area than a smaller
community square.

Note that this formulation is an extension of the gravity-based accessibility
statistics, in the same way that the gravity model is a specific case of
a destination-choice model \citep{Daly1982}. The extension allows us to consider
more attributes of the park or the indivdual beyond size and impedance. For example,
we can aggregate the number of tweets emanating from a park as a measure of the
park's attractiveness beyond its size, and the utility for the choice maker from 
Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

\section{Empirical Methodology}

\subsection{Data}

@@

This study used data available to the public from a variety of agencies and
data providers.
We obtained data on aggregate health outcomes from the Centers for Disease
Control and Prevention (CDC) for census tracts in five boroughs of New York City.
After removing tracts with missing population information, we have
Sexpr{nrow(tracts)} complete cases. Table \ref{tab:tractsdata} presents key
descriptive statistics for these data.

@@

summary_table(tracts@data, our_summary) %>%
  kable("latex", booktabs = T, escape = FALSE) %>%
  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
  group_rows("Physical Activity: Percent of tract reporting physical activity", 5, 8) %>%
  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
  group_rows("College: percent of adults with a college degree", 17, 20)
@
\end{table*}

@@

We also collected data on park geometry, limited to parks larger than 0.5 acres.

\begin{table*}
  \caption{Summary Statistics of Parks Data}
  \label{tab:parksdata}
<<parks_summary, results="asis", message=FALSE, warning=FALSE, eval = FALSE>>=
park_summary <- list(
  "size"     = tab_summary(parks$size),
  "tweets"   = tab_summary(parks$tweets)
)

summary_table(as_data_frame(parks), park_summary) %>%
  kable("latex", booktabs = T, escape = FALSE, col.names = c("")) %>%
  group_rows("Size: size of park in acres", 1, 4) %>%
  group_rows("Tweets: number of tweets emanating from park", 5, 8)
@
\end{table*}

@@

>>>>

The United States is currently facing an epidemic of obesity and chronic
diseases, which are non-communicable diseases of long duration and typically
slow development, including cardiovascular diseases, chronic respiratory
diseases, diabetes, stroke, joint and bone diseases, and cancer \citep{WHO2014}
Recent statistics suggest that approximately 300,000 premature deaths each year
can be attributed to chronic diseases \citep{WHO2014}.
According to \citet{Finkelstein2009}, obese Americans spend approximately
\$1,500 more in healthcare costs annually than Americans of
average weight, totaling \$147 billion in direct medical expenses nationally.

While a moderate amount of regular physical activity has been established as an
effective strategy for reducing and managing obesity and many of the
aforementioned chronic diseases \citet{CDC2009, Durstine2013},
\citet{Wolf2008} notes that “more than 50 percent of U.S. adults do not get enough
physical activity to provide health benefits; 24 percent are not active at all
in their leisure time. Activity decreases with age and sufficient activity is
less common among women than men, and among those with lower incomes
and less education” (p. 22). As the epidemiological transition from
infectious to chronic diseases is now complete in the developed world,
increasing physical activity has become a vital public health task in the 21st
century.

Until recently, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
According to \citet{Coutts2008} however, this trend is shifting as professionals
“have begun adopting an ecological paradigm, accepting that both individual and
environmental determinants play a role in health behavior. This new, arguably
revisited, public-health paradigm accounts not only for the compositional (who
you are) but also for the contextual (where you are) influences on physical
activity.” As professionals begin to operate from the assumption that the design
and configuration of the built environment can facilitate or inhibit physical
activity, they are increasingly looking to public spaces like parks and
greenways as key elements of the built environment that can support exercise
\citep{Bedimo-Rung2005a}.

In this paper, we present a holistic and flexible measurement for park
accessibility based in random utility theory. This measure can consider the continuous
distance to all parks in the region, weighted by the size of the park and other
amenities. We apply this measurement to study the link between park
accessibility and attractiveness and Census tract-level aggregate physical
activity participation and obesity rates in New York City, controlling for
spatial correlation and dependence. We find a positive relationship where the
least park-accessible tracts have an expected physical activity participation
rate 3.8 percent lower than the most accessible tracts, controlling for
socioeconomic characteristics. We also find that increased physical activity
rates lead to reduced aggregate obesity rates, but that when controlling for
physical activity rates, accessibility to parks does not reduce overall accessibility.
This suggests that to be an effective tool for addressing chronic diseases,
communities must do more than simply provide parks, but also incentivize more
citizens to use them.

\section{Literature Review}\label{sec:litreview}

Numerous previous studies have found relationships between aspects of the built
environment and aspects of physical, mental, social, and economic health. Many
of these findings are summarized in Table \ref{tab:parkbenefits}.
For the purposes of this study, it is clear that parks and other
vegetated areas support outdoor physical activity in ways distinct from other
urban environments \citep{Giles-Corti2005, Wells2007}.

\begin{table*}[ht]
  \caption{Commonly cited benefits of parks and greenspaces}
  \label{tab:parkbenefits}
\begin{tabular}{lp{11cm}}
\toprule
Category & Summary of Benefits\\
\midrule
Physical Health & Provide clean drinking water \citep{Benedict2006}\\
 & Promote faster healing in hospitals \citep{Akbari2001}\\
 & Reduce heat-related mortality \citep{Stone2006}\\
 & Reduce incidence of cardiovascular-related mortality \citep{Mitchell2008}\\
 & Improved air quality and related reductions in respiratory-related mortality \citep{Lovasi2008}\\
\midrule
Mental Health & Reduce stress and mental fatigue \citep{Wolch2014}\\
  & Reduce aggression \citep{Kuo2001a}\\
  & Enhance emotional and cognitive development\\
  & Improve behavioral outcomes in youth \\
\midrule
Social Health & Enhanced community aesthetics\\
 & Reduce crime \citep{Kuo2013,Wolfe2012}\\
 & Increase social interaction \citep{Sullivan2004}\\
\midrule
Economic Health & Provide ecosystem services\\
 & Increase residential property values and municipal property tax revenues\\
 & Attract more shoppers and increase economic activity to commercial districts \\
\bottomrule
\end{tabular}
\end{table*}

In spite of these generally positive findings, the evidence of a specific link
between green space and physical activity is somewhat mixed.
\citet{Wolf2008} report
residents self-identify as more active and attaining a higher quality of life when
a greater portion of the environment is greenspace.
\citet{Larson2016}
used self-reported scores on the Gallup-Healthways Wellbeing
Index (citation) to evaluate the relationship between different areas of wellbeing,
including physical health, and park quantity, quality, and accessibility in 44
U.S. cities. While the authors found positive associations between wellbeing
and park quality and accessibility, these relationships were not statistically
significant. A study by \citet{West2012} used park data from the
Trust for Public Land’s 2010 City Park Facts and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine relationships
between the density of parkland, parkland per capita, and levels of physical
activity and obesity for 67 metropolitan statistical areas in the U.S. The study
found a significant, positive association between park density and physical
activity and a significant, negative association between park density and
obesity.
Conversely, \citet{Richardson2012} examined the
relationship between urban greenspace and selected mortality rates, and though the authors
did not find a statistically significant relationship between the quantity of
urban greenspace and mortality caused by lung cancer, diabetes, heart disease or
car accidents, they did find that all-cause mortality was significantly higher in
greener cities.
In a meta-analysis of 20 peer reviewed journal articles
exploring the relationship between parks and objectively measured physical
activity, \citet{Bancroft2015} found that five studies reported a significant
positive association between the two, six studies produced mixed results, and
nine studies found no association at all.

While several of the aforementioned studies yielded results consistent with the
hypothesis that park access and use confer health benefits stemming from
increased physical activity and attendant decreases in obesity, all of the
studies -- except some included in the meta-analysis by \citet{Bancroft2015} -–
were conducted at the level of the city or metropolitan statistical area.
Metropolitan-level analyses do not capture the within-city variation in
accessibility to parks that exists. Some cities with large amounts of total
greenspace may nevertheless inequally distribute this space throughout the city
leading to areas with poor access. Conversely, a city with a smaller overall
proportion of greenspace may give all of its citizens better access to significant
greenspace. Considering access at the neighborhood level within a single city
may eliminate some regional or cultural fixed effects affecting
metropolitan-level analyses.

In a study of neighborhoods in New York City, \citet{Stark2014} found that a
higher proportion of park space in a neighborhood was associated with a lower
BMI for its residents. \citet{Pretty2005} showed the presence of trees and other
vegetation in outdoor environments is positively associated with physical
activity; further, \citet{Suminski2005} showed that adding vegetation to
existing sidewalks and trails increased local motivation to engage in physical
activity.

To evaluate the impacts of parks and greenspaces on health outcomes at a
sub-metropolitan level, one of the fundamental tasks involved is the selection
of a measure of accessibility. In a comprehensive meta-analysis of published
studies that measure active accessibility (accessibility using non-motorized
travel modes including walking and cycling), \citet{Vale2016} identified four
broad categories of studies based on the metric employed:
\begin{description}
  \item[Distance-based]{Account only for the Euclidean distance between origins
  and destinations}
  \item[Infrastructure-based]{Explicitly incorporate
  relevant transportation networks like roads and sidewalks to more accurately
  measure travel time and distance}
  \item[Gravity-based]{Incorporate travel impedance
  measures to model accessibility as a function of a destination’s attractiveness}
  \item[Walk-score]{Similar to Gravity-based, with the impedance measure
  explicitly designed around walking}
\end{description}

Importantly, \citeauthor{Vale2016} acknowledge that there is not yet a
consensus on the most appropriate accessibility metric to use in a given
setting, noting that ''ways to measure active accessibility are as varied as the
number of scholars the measure them.'' For example, infrastructure-based distances
can vary widely depending on which modes of travel infrastructure are considered.
Similarly, gravity-based measures may not always account for surrounding land
uses when considering the attractiveness or impedance of a destination.
In evaluating methodologies of measuring access to urban services, including
parks and greenspaces, \citet{Logan2017} note that existing and commonly used
approaches "often simplify their measure of proximity by using large areal
units and by imposing arbitrary distance thresholds," which often results in
access-poor populations being overlooked.

@@

A key theoretical outcome of random utility choice models is that the
consumer surplus, or the total value of all alternatives of the choice set, can
be obtained as the log-sum of the denominator of the choice probability equation
\citep{Small1981}:
\begin{equation}\label{eq:cs}
CS_i = \ln\left({\sum_{p \in J}\exp(V_{ip})}\right)
\end{equation}
There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary cutoff asserted by the researcher. This allows for the fact that some
people are more or less sensitive to distances, and that distance is a
continuous, and not a binary, phenomenon. Second, the random utility formulation
allows the researcher to include any attribute of the park; in this case, we
consider the size of the park as an element of accessibility. This suggests that
not all parks are equal, and that a large park such as Central Park on Manhattan
may provide health and activity benefits over a much larger area than a smaller
community square.

Note that this formulation is an extension of the gravity-based accessibility
statistics, in the same way that the gravity model is a specific case of
a destination-choice model \citep{Daly1982}. But the extension is meaningful, and
allows us to consider more attributes of the park or the indivdual beyond size
and impedance. For example, we can aggregate the number of tweets emanating from
a park as a measure of the park's attractiveness beyond its size, and the
utility for the choice maker from Equation \ref{eq:v_sd} then becomes
\begin{equation}\label{eq:v_sdt}
V_{ij} = size_j\lambda_s + distance_{ij}\lambda_d + tweets_j\lambda_t
\end{equation}

In spite of its advantages, logsum-based accessibilities have not received
as much application in the accessibility literature as distance-based or even
gravity-based measures. They are commonly used, however, in alternatives analyses of
transit infrastructure improvements \citep{DeJong2007}. A reason for this is
likely that such analyses are usually conducted in the context of a travel
demand model, where calibrated and multimodal logsums are readily available
\citep{Geurs2010}.

\section{Empirical Methodology}
In this section we describe an experiment with data for New York City, where
we compute a holistic logsum-based accessibility component and model the
relationship between this measure and physical activity rates, controlling for
spatial effects and socioeconomic factors.

\subsection{Data}
This study uses data available to the public from a variety of agencies and
data providers.

@@

The primary dataset is a geographic polygons shapefile of Census tracts in
New York City. We appended relevant sociodemographic data for each tract level
from the American Community Survey 2011-2015 5-year estimates. Data on physical
activity participation rate and obesity rates for each tract are available from
the Centers for Disease Control and Prevention’s 500 Cities Project.
After removing tracts with missing population information or outlying and unusual
characteristics, we have Sexpr{nrow(tracts)} complete cases. Table
\ref{tab:tractsdata} presents key descriptive statistics for these data.

@@

summary_table(tracts@data, our_summary) %>%
  kable("latex", booktabs = T, escape = FALSE) %>%
  group_rows("Obese: percent of tract which is obese", 1, 4) %>%
  group_rows("Physical Activity: percent of tract reporting physical activity", 5, 8) %>%
  group_rows("Population Density: individuals per square mile", 9, 12)  %>%
  group_rows("Full Time: percent of adults who work full time", 13, 16) %>%
  group_rows("College: percent of adults with a college degree", 17, 20)
@
\end{table*}

In a destination choice framework, the tracts represent the "origins" and
"destinations" are parks in New York City.
We retrieved a shapefile of public parks and greenspaces within New York City’s
municipal boundaries and checked it for accuracy and relevance. Upon inspection,
we removed several facilities that do not qualify as publicly accessible green
space, such as Yankee Stadium and its surrounding parking lots. We also removed
parks of less than half an acre in size, as these appear to be predominantly
vacant lots or parking strips rather than legitimate public green space.

@@

\begin{table*}
  \caption{Summary Statistics of Parks Data}
  \label{tab:parksdata}
<<parks_summary, results="asis", message=FALSE, warning=FALSE, cache = FALSE>>=
park_summary <- list(
  "size"     = tab_summary(parks$size),
  "tweets"   = tab_summary(parks$tweets)
)

summary_table(as_data_frame(parks), park_summary) %>%
  kable("latex", booktabs = T, escape = FALSE, col.names = c("Parks"))
@
\end{table*}

@@

<<obesity_models>>=
obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ . + log_physact),
                           data = tracts@data, listw = W, zero.policy = TRUE,
                           type = "mixed")
obese_access_sdm <- update(obese_base_sdm, . ~ . + access_ls + access_ls * log_physact)
obese_tweets_sdm <- update(obese_base_sdm, . ~ . + tweets_ls + access_ls * log_physact)
obesity_models <- list("Obesity" = obese_base_sdm, "Obesity Access" = obese_access_sdm, 
                       "Obesity Tweets + Access" = obese_tweets_sdm)
@

<<obese_impacts>>=
obese_impacts <- obesity_models %>%
  lapply(extract_impacts) %>%
  bind_rows(.id = "model") %>% ungroup() %>%
  rename(submodel = effect)

dwplot(bind_rows(physact_impacts, obese_impacts) %>% 
                 filter(term %in% c("log_physact", "access_ls", "tweets_ls")) %>%
                 filter(submodel == "Total")) +
  theme_bw()

obese_impacts %>%
  transmute(
    model, term, submodel, estimate,
    output = paste(round(`p-val`, 5),
                   gtools::stars.pval(`p-val`))
  ) %>%
  filter(term %in% c("log_physact", "access_ls", "tweets_ls"), submodel == "Total")
@
